---
title: "Matching project"
author: "Nadir Nibras"
date: "January 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "C:/Users/nadir/Desktop/Matching/BKN 599 Advanced Data Analysis")
```


# Matching project for evaluating effects of treatment dosage
clear workspace variables
clear command window (same as ctrl+L. )
close all plots
```{r}
rm(list = ls()); 
cat("\014")  
graphics.off() 
```

```{r}
library(tidyverse)
library(Matching)
library(cobalt)
library(onehot)
library(pander)
library(gridExtra)
```

### Thresholds
```{r}
Lthreshold= 15
Hthreshold= 30
```

### Read files
```{r}
setwd("C:/Users/nadir/Desktop/Matching/BKN 599 Advanced Data Analysis")
evalData=read.csv("00042_evalDataset_notext.csv")
patChar= read.csv("PatientCharacteristics_withDose.csv")
rawd= read.csv("Raw Data_Included Subjects Only_mod.csv", sep="\t") # neural values
library(readr)


genweight= c(0.97736335, 0.68557564, 0.11435119, 0.55160536, 0.95083754, 0.04852164, 0.14841358, 0.19860284, 0.79674611, 0.79042807, 0.62237024, 0.01534745, 0.32553671, 0.19962585, 0.64700365, 0.15999446, 0.12968842, 0.02935539, 0.45110192, 0.16812796, 0.51002147, 0.81456330, 0.59208731, 0.20140278, 0.26642759, 0.15331324, 0.34255925, 0.37867908, 0.01936102, 0.06611066, 0.47779205, 0.38559177, 0.88105505, 0.94637622, 0.11744363, 0.29967114, 0.29017548, 0.37624276, 0.83526589, 0.84839756, 0.46941722, 0.21737904, 0.40772396, 0.28270889, 0.78696707, 0.75454488, 0.08591095, 0.56542856, 0.79789390, 0.90011766, 0.50598825, 0.31844054, 0.84565483, 0.59102811, 0.51581391, 0.11314781, 0.70154351)



```

### Coding all -megatove values as missing data
```{r}
evalData[evalData == -99] <- NA;
patChar[(patChar ==-99)|(patChar ==-7)|(patChar ==-8)|(patChar ==-9)] <- NA;
rawd[rawd ==-1] <- NA;
```

#### Setting key for merging datasets 
```{r}
patChar=rename(patChar, pid= PID )
rawd=rename(rawd, pid= Study_id )
```

### Examining Matching features from EvalData visit 1 scores
```{r}
# evalData.cnames= data.frame(colnames(evalData))
# pc.cnames= data.frame(colnames(patChar))
# rawd.cnames= data.frame(colnames(rawd))
# evalData1=evalData1[complete.cases(evalData1),]
# evalD.correlations= data.frame(cor(evalData1[-c(1,2)]))
```

# Processing Raw/Neural Data set
```{r}
rawd$multlesions= as.integer(rawd$A2x!="\\N")
rawd= rawd[c(1,13,32, 107, 130, 143)];
rawd$old_stroke= as.integer(as.factor(rawd$old_stroke))-1
rawd$leukoaraiosis= as.integer(as.factor(rawd$leukoaraiosis))-1
rawd$A1EstVol= log2(rawd$A1EstVol+1)
rawd=rename(rawd, log2.A1EstVol= A1EstVol ) #renaming column name
```

# Remove sparse features/features with no use
```{r}
# evalData = evalData[c(5,6,120,128,204,246,260,277,367,368,371,372)]
evalData = evalData[c(5,6,107,120,128,204,246,260,277,367,368,371,372)]
# Note: Leaving out FAS scores because many (18) missed values for visit 1 stats
patChar= patChar[c(3,6,7,12,15:21,27,29:32)]

#removing Treatment type
patChar= patChar[-c(3)]
```


### Merge data
```{r}
dataMerged= merge(evalData,patChar)
dataMerged= merge(dataMerged, rawd)
```


### Extract visit datas
```{r}
dataVisit1= subset(dataMerged, visitnum==1)                       
dataVisit2= subset(dataMerged, visitnum==2)       
rm(dataMerged)
```


### Removing subjects with missing values
```{r}
dataVisit1= dataVisit1[complete.cases(dataVisit1),]
dataVisit2= dataVisit2[complete.cases(dataVisit2),]
```


### Removing subjects who don't attend visits 1 and 2
```{r}
temp1= data.frame(dataVisit1$pid); 
temp2= data.frame(dataVisit2$pid); 

names(temp1)[1]<-paste("pid");   #renaming column name to "pid"
names(temp2)[1]<-paste("pid");   #renaming column name to "pid"

temp= merge(temp1,temp2)

dataVisit1= merge(temp,dataVisit1)
dataVisit2= merge(temp,dataVisit2)

rm(temp,temp1,temp2)
```


### Creating dataset for matching
```{r}
data12= dataVisit1
```


### One hot encoding
```{r}
# Site code
sitecode=data.frame(as.factor(data12$siteCode)) 
names(sitecode)[1]<-paste("Site.code")
data12 = cbind(data12,data.frame(predict(onehot(sitecode),sitecode)))

# Predominant areas affected
predominant=data.frame(as.factor(data12$predominant)) 
names(predominant)[1]<-paste("predominant")
data12 = cbind(data12,data.frame(predict(onehot(predominant,  max_levels = 13),predominant)))
# combing ACs and MCs into same categories

data12$predominant.AC= ifelse((data12$predominant.AC1==1)|(data12$predominant.AC2==1)|(data12$predominant.AC3==1),1,0)
data12$predominant.MC=ifelse((data12$predominant.MC2==1)|(data12$predominant.MC3==1)|(data12$predominant.MC5==1),1,0)

# Ethnicity- lack of both indicates declined to report
data12$hispanic.latino= ifelse(data12$DEMOEthnicity==1, 1, 0)
data12$nonhispanic.latino= ifelse(data12$DEMOEthnicity==2, 1, 0)

# Race
# Few subjects are indegenous/pacufic islanders
data12$race.other= ifelse((data12$DEMORace==1)|(data12$DEMORace==3)|(data12$DEMORace==6), 1, 0) 
data12$race.asian= ifelse(data12$DEMORace==2, 1, 0)
data12$race.africanam= ifelse(data12$DEMORace== 4, 1, 0)
data12$race.caucasian= ifelse(data12$DEMORace== 5, 1, 0)


# Education
# under/some high school 
data12$edu.noHSdegree= ifelse((data12$DEMOedu==1)|(data12$DEMOedu==2), 1, 0) 
data12$edu.finishHS= ifelse(data12$DEMOedu==3, 1, 0)
data12$edu.someCollege= ifelse(data12$DEMOedu== 4, 1, 0)
# Bachelors, Masters and PhD grads piled into 1 category
data12$edu.finishCollege= ifelse((data12$DEMOedu== 5)|(data12$DEMOedu== 6)|(data12$DEMOedu== 7), 1, 0)

# Stroke location
data12$stroke.loc.RightH= ifelse(data12$CSstrkloc==1, 1, 0)
data12$stroke.loc.LeftH= ifelse(data12$CSstrkloc==2, 1, 0)
data12$stroke.loc.cerebellar= ifelse(data12$CSstrkloc==3, 1, 0)
data12$stroke.loc.brainstem= ifelse(data12$CSstrkloc== 4, 1, 0)

# Stroke type
data12$stroke.type.ISCHEMIC.wo.hemorrhagic.conversion= ifelse(data12$CSstrktype==1, 1, 0)
data12$stroke.type.ISCHEMIC.with.hemorrhagic.conversion= ifelse(data12$CSstrktype==2, 1, 0)
data12$stroke.type.intraparenchymal.HEMORRHAGIC= ifelse(data12$CSstrktype==3, 1, 0)
data12$stroke.type.other= ifelse(data12$CSstrkloc== 4, 1, 0)

#Remove extra columns and extra variables
data12$DEMOedu=NULL
data12$DEMOEthnicity=NULL
data12$DEMORace=NULL
data12$CSstrkloc=NULL
data12$CSstrktype=NULL
data12$predominant=NULL
data12$siteCode=NULL
data12$visitnum= NULL
data12$predominant.AC1=NULL
data12$predominant.AC2=NULL
data12$predominant.AC3=NULL
data12$predominant.MC2=NULL
data12$predominant.MC3=NULL
data12$predominant.MC5=NULL
rm(sitecode,predominant)
```


### Changing table to numeric
```{r}
for (i in 2:ncol(data12)){
  data12[c(i)]= as.numeric(unlist(data12[c(i)]))
}
```


### Creating visit2 values, and recovery parameters from differences in visit2 and vist 1
```{r}
data12$SIS_V2= dataVisit2$SIS_hand
data12$fugm_V2= dataVisit2$ufugm
data12$logwmftMA_V2= dataVisit2$log_mean_time_MA_PA

data12$sishand_recovery= dataVisit2$SIS_hand-dataVisit1$SIS_hand
data12$fugm_recovery= dataVisit2$ufugm-dataVisit1$ufugm
data12$logwmftMA_recovery= dataVisit2$log_mean_time_MA_PA-dataVisit1$log_mean_time_MA_PA
```


# Note: Low Dose group is the treated group, and High Dose is the control group for matching
### creating binary variable indicating whether subject belongs in Low Dose or High Dose group  
```{r}
data12$treat=ifelse(data12$dose_hours<=27, 1, 0)
```


### This line removes all subjects between 16-29 hours of dosage and changes starting weights
```{r}
data12=data12[data12$dose_hours>=Hthreshold | data12$dose_hours<=Lthreshold,]

genweight = c(0.830426260, 0.724817864, 0.651582563, 0.297741502, 0.830889340, 0.087003014,
 0.638399491, 0.121175618, 0.969273674, 0.785379687, 0.190612974, 0.816637110,
 0.880788106, 0.631580992, 0.599959964, 0.437396421, 0.300651108, 0.942747991,
 0.034287552, 0.447049238, 0.943430585, 0.214020309, 0.831556053, 0.204920316,
 0.834063460, 0.265514404, 0.025610668, 0.149471313, 0.796257628, 0.333164340,
 0.744276975, 0.136378485, 0.624898450, 0.736891277, 0.045765780, 0.510092046,
 0.742510688, 0.958445071, 0.628410029, 0.566133301, 0.190156881, 0.155908728,
 0.191743071, 0.058801373, 0.342044210, 0.396892759, 0.122116592, 0.107519303,
 0.420979936, 0.087684545, 0.404056359, 0.005771743, 0.650855758, 0.224358681,
 0.666497053, 0.120712778, 0.875230230)

```



### Creating 2 dataframes separatomg low dose and high dose (not used)
```{r}
highdosegroup= subset(data12, treat==0)
lowdosegroup= subset(data12, treat==1)
```


### Propensity Score Matching code starts here

creates a matrix from columns/features
```{r}
X= data.matrix(data12[,c(2:12,14:58)])
```


### Calculating the propensity score and adding it to data set
Generalized linear models are fit using the glm( ) function.
treat ~ X means treat is the response of all other variables
treat is a binary factor and Family objects provide a convenient way to specify the details of the models used by functions such as glm
family= binomial() specifies that Logistic Regression is to be performed


```{r}
model <- glm(treat ~ X,    family = binomial(), data = data12 )
data12$propensity.score= model$fitted.values
```


### Matching and estimating effects of causality using "Match" package
"Match" implements a variety of algorithms for multivariate matching including propensity score, Mahalanobis and inverse variance matching. The function is intended to be used in conjunction with the MatchBalance function which determines the extent to which Match has been able to achieve covariate balance
- Y=y
A vector containing the outcome of interest. Missing values are not allowed. An outcome vector is not required because the matches generated will be the same regardless of the outcomes. Of course, without any outcomes no causal effect estimates will be produced, only a matched dataset.
- X= model$fitted.values
A matrix containing the variables we wish to match on. This matrix may contain the actual observed covariates or the propensity score or a combination of both. All columns of this matrix must have positive variance or Match will return an error
Note: In our case it is the propensity score (fitted.values are predicted values for log regr.)
- estimand = "ATT"
A character string for the estimand. The default estimand is "ATT", the sample average treatment effect for the treated. "ATE" is the sample average treatment effect, and "ATC" is the sample average treatment effect for the controls.
- M = 1
A scalar for the number of matches which should be found. The default is one-to-one matching. Also see the ties option

-replace = TRUE
Matching with replacement


### Genetic matching 
This function finds optimal balance using multivariate matching where a genetic search algorithm determines the weight each covariate is given. Balance is determined by examining cumulative probability distribution functions of a variety of standardized statistics. By default, these statistics include t-tests and Kolmogorov-Smirnov tests. A variety of descriptive statistics based on empirical-QQ (eQQ) plots can also be used or any user provided measure of balance. The statistics are not used to conduct formal hypothesis tests, because no measure of balance is a monotonic function of bias and because balance should be maximized without limit. The object returned by GenMatch can be supplied to the Match function (via the Weight.matrix option) to obtain causal estimates. GenMatch uses **GENoud** to perform the genetic search. 
- Tr= data12$treat
A vector indicating the observations which are in the treatment regime and those which are not. This can either be a logical vector or a real vector where 0 denotes control and 1 denotes treatment.
- X=X
A matrix containing the variables we wish to match on. 
#####This matrix may contain the actual observed covariates or the propensity score or a combination of both.
- pop.size= 10000
Population Size. This is the number of individuals genoud uses to solve the optimization problem. The theorems proving that genetic algorithms find good solutions are asymptotic in population size. Therefore, it is important that this value not be small. See genoud for more details.
-**GENoud**- GENetic Optimization Using Derivatives
Genoud is a function that combines evolutionary search algorithms with derivative-based (Newton or quasi-Newton) methods to solve difficult optimization problems.


# Adding propensity score to feature matrix used for for genetic matching 

```{r}
X1 = cbind(X,data12$propensity.score)
```

# Doing the actual genetic matching 
```{r}
minimump= c()

set.seed(1)

#Generating matches
gen1 = GenMatch(Tr= data12$treat, 
                starting.values = genweight,
                X=X1, pop.size=1000, replace=TRUE, ties= TRUE, max.weight=1)

# Doing the actual matching
mgen1=Match(Y= data12$sishand_recovery,BiasAdjust = TRUE, estimand= "ATT", Tr=data12$treat, X=X1, Weight.matrix = gen1)

```

## Assessing balance, checking minimum p value
Note that "MatchBalance" is used to determine if Match was successful in achieving balance on the observed covariates.

-treat~SIS_hand+ ...
This formula does not estimate any model. The formula is simply an efficient way to use the R modeling language to list the variables we wish to obtain univariate balance statistics for. The dependent variable in the formula is usually the treatment indicator

-nboots= 1000
The number of bootstrap samples to be run. If zero, no bootstraps are done. Bootstrapping is highly recommended because the bootstrapped Kolmogorov-Smirnov test provides correct coverage even when the distributions being compared are not continuous. At least 500 nboots (preferably 1000) are recommended for publication quality p-values.

## Note that commenting out propensity score in the code below doesn't mean that weights for propensity score wasn't used for creating matches during genetic matching. It just means that propensity scores are not looked at when evaluating overall balance acheived after Matching
```{r}
#using 57 features
balanceA = MatchBalance(treat~ ave_CAHM + CHAMchallenge + SIS_hand + ufugm + 
               NIHtot + log_mean_time_MA_PA + log_mean_time_LA_PA + EQ_index +
                grip_MA +grip_LA +RNLIadj+
                 
               CSgender+ concordance +  onset_to_rand + CStpa + age_at_rand +
               FCtotalOT+ log2.A1EstVol+old_stroke +DEMOsmoke+ leukoaraiosis+ multlesions+

               Site.code.1+ Site.code.2+ Site.code.3+Site.code.4+
               Site.code.5+ Site.code.6+ Site.code.7+ 
                 
                 predominant.AC+ predominant.ACH +predominant.basilar+ predominant.IB+
               predominant.LS+ predominant.MC+ predominant.PCA+
               predominant.vertebral +predominant.THP+ 
               
                hispanic.latino+nonhispanic.latino+
               race.asian+ race.africanam+ race.caucasian + race.other+ 
                 
                 edu.noHSdegree +edu.finishHS +edu.someCollege + edu.finishCollege+ 
              
                 stroke.loc.RightH+ stroke.loc.LeftH +
                 stroke.loc.cerebellar +stroke.loc.brainstem+
               
                 stroke.type.ISCHEMIC.wo.hemorrhagic.conversion+
               stroke.type.ISCHEMIC.with.hemorrhagic.conversion +
               stroke.type.intraparenchymal.HEMORRHAGIC + stroke.type.other
            + propensity.score
           ,
               
               data= data12, match.out= mgen1, nboots=2000)

minimump=c(minimump, balanceA$AMsmallest.p.value) 

print(balanceA$AMsmallest.p.value)
```

####Summary.Match of the model will tell us about Causal relationship after balancing
```{r}
mgen1=Match(Y= data12$sishand_recovery,BiasAdjust = TRUE, estimand= "ATT", Tr=data12$treat, X=X1, Weight.matrix = gen1)
summary.Match(mgen1, full = TRUE)
```


### Checking number of unique subjects after genetic matching
```{r}
length(unique(mgen1$index.treated))
length(unique(mgen1$index.control))
```


#Additional method of checking balance (STROKE.TYPE.OTHER not showing??)
```{r}
(bal.tab(mgen1, treat~ ave_CAHM + CHAMchallenge + SIS_hand + ufugm + 
               NIHtot + log_mean_time_MA_PA + log_mean_time_LA_PA + EQ_index +
                grip_MA +grip_LA +RNLIadj+

               CSgender+ concordance +  onset_to_rand + CStpa + age_at_rand +
               FCtotalOT+ log2.A1EstVol+old_stroke +DEMOsmoke+ leukoaraiosis+ multlesions+

               Site.code.1+ Site.code.2+ Site.code.3+Site.code.4+
               Site.code.5+ Site.code.6+ Site.code.7+ 
           
               predominant.AC+  predominant.ACH +predominant.basilar+ predominant.IB+
               predominant.LS+ predominant.MC+ predominant.PCA+  predominant.vertebral +predominant.THP+ 
               
               hispanic.latino+nonhispanic.latino+
           
               race.asian+ race.africanam+ race.caucasian + race.other+ 
                 
                 edu.noHSdegree +edu.finishHS +edu.someCollege + edu.finishCollege+ 
              
                 stroke.loc.RightH+ stroke.loc.LeftH +
                 stroke.loc.cerebellar +stroke.loc.brainstem+
           
               stroke.type.ISCHEMIC.wo.hemorrhagic.conversion+
               stroke.type.ISCHEMIC.with.hemorrhagic.conversion +
               stroke.type.intraparenchymal.HEMORRHAGIC + stroke.type.other+
           
              propensity.score,

         disp.v.ratio = TRUE, data = data12,abs= TRUE, m.threshold =0.25))
```


# Plotting distributions of continuous variables before and after matching

```{r}
columns = colnames(data12)
for (i in c(2:12,14,16,18,21)){
  string= columns[i]
  
  b= max(data12[,i])/25
  
  plot1<-ggplot(data12, aes(eval(parse(text = string)))) +
    geom_histogram(data= data12[mgen1$index.treated,],fill="red",alpha =0.2, binwidth=b)+
    geom_histogram(data= data12[-mgen1$index.treated,],fill="blue",alpha =0.2, binwidth=b)+
    ggtitle("Low Dose(Red) vs Unmatched High Dose (Blue)")+
    theme(plot.title = element_text(face="bold", size=10)) +
    xlab(paste(string, "histogram"))+
    coord_cartesian(ylim=c(0,35))+
    xlim(min(data12[,i]),max(data12[,i]))
  
  plot2<- ggplot(data12, aes(eval(parse(text = string)))) +
    geom_histogram(data= data12[mgen1$index.treated,],fill="red",alpha =0.2, binwidth=b)+
    geom_histogram(data= data12[mgen1$index.control,],fill="blue",alpha =0.2, binwidth=b)+
    ggtitle("Low Dose(Red) vs Matched High Dose (Blue)")+
    theme(plot.title = element_text( face="bold", size=10)) +
    xlab(paste(string, "histogram"))+
    coord_cartesian(ylim=c(0,35))+
    xlim(min(data12[,i]),max(data12[,i]))
  
   plot3<-ggplot(data12, aes(eval(parse(text = string)))) +
    geom_density(data = data12[mgen1$index.treated,],fill = "red", alpha = 0.2)+
    geom_density(data = data12[-mgen1$index.treated,],fill = "blue", alpha = 0.2)+
    ggtitle("Low Dose(Red) vs Unmatched High Dose (Blue)")+
    theme(plot.title = element_text(face="bold", size=10)) +
    xlab(paste(string, "density"))+
    xlim(min(data12[,i]),max(data12[,i]))
   
   plot4<-ggplot(data12, aes(eval(parse(text = string)))) +
    geom_density(data = data12[mgen1$index.treated,],fill = "red", alpha = 0.2)+
    geom_density(data = data12[mgen1$index.control,],fill = "blue", alpha = 0.2)+
    ggtitle("Low Dose(Red) vs Matched High Dose (Blue)")+
    theme(plot.title = element_text(face="bold", size=10)) +
    xlab(paste(string, "density"))+
    xlim(min(data12[,i]),max(data12[,i]))
  
  grid.arrange(plot1, plot2, plot3, plot4, ncol=2, nrow=2)
}

```


# Plotting distributions of binary variables before and after matching


```{r}
for (i in c(15,17,19,20,22:54)){
  string= columns[i]
  
  plot1<-ggplot(data12, aes(eval(parse(text = string)))) +
    geom_histogram(data= data12[mgen1$index.treated,],fill="red",alpha =0.2, binwidth = 0.5)+
    geom_histogram(data= data12[-mgen1$index.treated,],fill="blue",alpha =0.2, binwidth = 0.5)+
    ggtitle("Low Dose(Red) vs Unmatched High Dose(Blue)")+
    theme(plot.title = element_text(face="bold", size=10)) +
    xlab(paste(string, "histogram"))+
    coord_cartesian(ylim=c(0,150))  
  
  plot2<- ggplot(data12, aes(eval(parse(text = string)))) +
    geom_histogram(data= data12[mgen1$index.treated,],fill="red",alpha =0.2, binwidth = 0.5)+
    geom_histogram(data= data12[mgen1$index.control,],fill="blue",alpha =0.2, binwidth = 0.5)+
    ggtitle("Low Dose(Red) vs Mmatched High Dose(Blue)")+
    theme(plot.title = element_text(face="bold", size=10)) +
    xlab(paste(string, "histogram"))+
    coord_cartesian(ylim=c(0,150))
  
  grid.arrange(plot1, plot2, ncol=2)
}
```


## Extra stuff

```{r}
 dose_hours= c(data12$dose_hours[mgen1$index.treated],data12$dose_hours[mgen1$index.control])
 treat= c(data12$treat[mgen1$index.treated],data12$treat[mgen1$index.control])
 SIS_Hand.recovery= c(data12$sishand_recovery[mgen1$index.treated],data12$sishand_recovery[mgen1$index.control])
 WMFT.recovery= c(data12$logwmftMA_recovery[mgen1$index.treated],data12$logwmftMA_recovery[mgen1$index.control])

datap= data.frame(dose_hours, treat, SIS_Hand.recovery,WMFT.recovery)

ggplot(datap,aes(x=dose_hours,y=SIS_Hand.recovery))+
  geom_jitter(alpha=0.7)+
  geom_smooth(method='lm')+
  xlab("Dose hours")+
  ylab("SIS Hand score change: Visit 2- Visit1")
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
  
ggplot(datap,aes(x=treat,y=SIS_Hand.recovery))+
  geom_point(alpha=0.3)+
  geom_smooth(method='lm')+
  xlab("Low Dose (1) vs High Dose (0)")+
  ylab("SIS Hand score change: Visit 2- Visit1")
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
  
ggplot(datap,aes(x=dose_hours,y=WMFT.recovery))+
  geom_point(alpha=0.7)+
  geom_smooth(method='lm')+
  xlab("Dose hours")+
  ylab("log(WMFT) change: Visit 2- Visit1")
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
  
#Regression analysis on SIS_Hand score
linearm=  lm(datap$SIS_Hand.recovery~ datap$treat)
pander(summary(linearm))
```

